---
title: "IC10, Owen, 5714"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## QUESTIONS

Q1. The factorial of a non-negative integer n is defined to be 
n!=n*(n-1)*(n-2)…*2*1,
so for example, 4!=4*3*2*1=24.  Create a function that takes a non-negative integer as the argument and returns the factorial of the integer.  (Hint: you can use a while loop, but there are many ways to do this).  What is the value of 10! ? 

```{r}
factorialint <- function(n){
  f=1;
  for (i in 1:n){
    f=f*i
  }
  return(f)
}
```

```{r}
factorialint(5)
5*4*3*2*1

factorialint(10)
10*9*8*7*6*5*4*3*2*1

```

Q2. The formula for converting a temperature in Fahrenheit (F) to Celsius (C) is:
C=(5/9)*(F-32)
Write a function that converts a Fahrenheit temperature to Celsius. Use this function to create a data frame containing Fahrenheit values 30, 31, 32, up to 100, and the corresponding temperatures in Celsius.

```{r}
FtoC <- function(F){
  C=(5/9)*(F-32)
  return(C)
}
```

```{r}
FtoC(32)
FtoC(95)
FtoC(65)
```

Q3. Obtain a root for the following function with the Newton-Raphson method:
f(x)=5x^3-4x^2+12x – 7
(Hint: Implement the Newton-Raphson function given in the session 10 slides).
```{r}
newton.raphson <- function(f,fprime,x0,thresh){
  myabsdiff <- Inf
  xold      <- x0
  while(myabsdiff>thresh){ # have we converged yet? If no, move;
    xnew      <- xold-f(xold)/(fprime(xold))
    myabsdiff <- abs(xnew-xold)
    xold      <- xnew
    }
  return(xnew)
}


```

```{r}
myf <- function(x){5*x^3+4*x^2+12*x-7}
myfprime <- function(x){15*x^2 +8*x +12}
r = newton.raphson(myf,myfprime,x0=0,thresh=0.0001)
r
myf(r)
curve(myf,-10,10)
```

